### 2025-05-14 manuscript/manuscript.Rmd
change manuscript format to Fisheries Research

bibliography      : ["citations.bib", "r-references.bib"]
csl               : "cjfas.csl"

@@

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
library(tidyverse)
library(stargazer)
library(kableExtra)
library(gridExtra)
library(nplyr)
library(patchwork)
library(ggh4x)
library(cowplot)
library(gridExtra)
library(grid)
library(scales)
library(here)

# caching for quick knitting.  
#    set cache = FALSE to not cache
#    set cache.extra to a new value to remove all caches and re-evaluate
knitr::opts_chunk$set(cache.extra = 4, cache = FALSE)

@@

```{r figures, include = FALSE}
source(here("code", "analysis", "results_tables.R"), local = knitr::knit_global())
source(here("code", "analysis", "results_figures.R"), local = knitr::knit_global())
```

@@

# Introduction

@@

# Methods

This section is organized as follows: 
we first describe the CKMR kinship probabilities for protogynous hermaphrodites, then
present the different components of the simulation study to test this model: the (individual-based) 
operating model (OM) to simulate 'reality', the sampling model to generate 'observed' data from 
the OM, and finally the integrated fisheries stock assessment estimation model (EM) that uses
the 'observed' data and the derived CKMR kinship probabilities to estimate a set of parameters 
that allow generating 'estimated' abundance which can then be compared to the 'true' abundance 
generated by the OM. For each simulation component, we describe the configurations or scenarios 
we used to test the performance of our CKMR estimators (Figure \@ref(fig:simoverview-plot)).

## CKMR Model

@@

The ability to identify---or at least infer---the sex of a sampled individual's parent is crucial for 
applying the model, and in our model this is accomplished by sequencing the fish's
mitochondrial DNA (mtDNA) which is inherited from the female parent. For the purposes
of the simulation analyses, we assumed that different mtDNA haplotypes were equifrequent,
so that we were able to parameterize the mtDNA haplotype frequencies simply by $H_\mathrm{mt}$---the
probability that two randomly sampled mtDNAs from the population
are different in sequence. This quantity is also commonly called the mtDNA
haplotype diversity [@avise2012molecular], and parameterizing the simulations in terms of it allowed
us to avoid simulating mtDNA sequences
in all simulated individuals and also simplified the CKMR likelihood used to analyze the simulations as described
in the Supplement. An $\Hmt$ of 1 would indicate a 100% certainty
that two HSPs with the same mtDNA sequence shared a mother, while any value <1 means there is
a chance (which equals $1 - \Hmt$) that their mtDNA sequence is the same even though they did not
share a mother. We incorporate uncertainty due to $\Hmt$ into our CKMR model
equations (Equations A8, and A14-A17) and in the simulation, we explore the effects of 
misspecifying both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\Hmt$ in the estimation model (see below).

## Model Species

@@

## Operating Model

The OM is a discrete-time, individual-based model (IBM) implemented in R based on the 'fishSim' 
package [@R-fishSim]. We modified the package functions to accommodate 
our study needs (e.g., by adding the ability to model protogynous hermaphrodite 
life histories) and rewrote the slower-running R functions in C++ using the 'Rcpp' 
package [@eddelbuettel2013seamless] to increase computational efficiency. The basic
annual sequence, modeled loosely on gag life history,
was that sex transition and then spawning occurred at the beginning
of the year, followed by mortality, CKMR sampling, and age incrementation (Figure \@ref(fig:flowchart-plot)).
For computational efficiency, the simulation starts at age 1, given that most age-0 gag
are not available to the fishery and suffer high rates of natural mortality.

Simulated reproductive value for females and males of age $a$, denoted as  $\mathrm{RV}_{a,F}$ and
$\mathrm{RV}_{a,M}$, respectively, were used to produce age-1 offspring and assign parentage.
$\mathrm{RV}_{a,F}$ represents the age-specific expected number 
of age-1 offspring produced per female. This was modeled as a power function 
(Figure \@ref(fig:RC-plot), panel A) derived 
from female gag batch fecundity-at-age data, and scaled to achieve a stable population in 
the absence of fishing. To help produce recruitment variability, $\mathrm{RV}_{a,F}$ 
was randomly scaled up or down annually by sampling a reproductive value, 
$\mathrm{RV}_{a,t,F,b}$ for each breeding group, $b$, from a $\mathrm{Uniform}(0,1)$ 
distribution and rescaling such that $\sum_a{\mathrm{RV}_{a,F}e^{\eta_t}} = \sum_a{\sum_b\mathrm{RV}_{a,t,F,b}}$.
Here, $\eta_t$ represents the annual scaling factor, ensuring that the total RV 
across all ages and breeding groups remains consistent.
The realized number of age-1 offspring in a given year $t$
for a given female of age $a$ in breeding group $b$ was a Poisson random draw from 
$\mathrm{RV}_{a,t,F,b}$. All males were considered mature, and $\mathrm{RV}_M$ 
was used as the age-specific probability of being the father
to any age-1 offspring produced by females in the male's breeding group. The true shape and 
parameter values of $\mathrm{RV}_M$ are unknown. Here, we modeled it as a power function, and in 
the base OM its exponent was set equal to that of $\mathrm{RV}_F$ (Figure \@ref(fig:RC-plot), panel A), 
which corresponds to assuming age-specific $\mathrm{RV}_M$ equals that of the females (the current
assumption of the gag stock assessment). Mating was assumed to 
be fully random within breeding groups.

@@

The simulation was run for 119 years, starting with an unfished population and 
ending with a highly depleted population. For each OM scenario (see below), 40
independent replicates of the population were simulated, providing 40
unique whole-population pedigrees.

### OM Scenarios

We simulated four alternative versions of 'reality' in our OM to explore the
effect of uncertainty about male contribution to reproductive success: in addition to
the base case, one scenario included sperm limitation and two scenarios
simulated (age-specific) male reproductive value functions ($\mathrm{RV}_M^\mathrm{OM}$)
that differed from those of the females ($\mathrm{RV}_M^\mathrm{EM}$). In the base OM, male 
and female reproductive value were equal, corresponding 
to what is currently assumed in the gag stock assessment, but in two alternative 
scenarios, we specified $\mathrm{RV}_M^\mathrm{OM}$ functions that increased more or less steeply
from $\mathrm{RV}_F^\mathrm{OM}$ (Figure \@ref(fig:RC-plot), panel A).

The true $\mathrm{RV}_M$ in protogynous hermaphrodites, including gag,
is unknown. It is assumed that sperm limitation exists but it has not been
measured. However, it has been shown that skipped spawning is relatively common in female gag,
and this is thought to be linked to low male abundance [@lowerre2022spawning].
In the base OM configuration, declining male abundance had no effect (meaning all eggs produced
by females will get fertilized, regardless of sex ratio). Alternately, we simulated the effect of
sperm limitation as an increased
likelihood of a female not producing any offspring in a given spawning season as the sex ratio
declines. Note that a female skipping spawning was independent between years and thus there was no systematic
skipped spawning such as alternate year spawning. The relationship between the 
current proportion of males in a spawning group relative to unfished levels, $p_M$,
and the probability of females in that spawning group skipping spawning was
modeled as an exponential function of the form $1-\exp(-{\tau p_M})$ (Figure
\@ref(fig:RC-plot), panel B), where $\tau$ controls the slope of the curve. We used a sigmoid
function to map probability of spawning to age, such that skipped spawning was more likely
in younger females. Thus, skipped spawning serves as a de facto modifier of the vector 
describing the age-specific probability of a female producing offspring in a given year.

## Sampling Model
Three types of observations were generated from the OM for use in the
integrated assessment model: CKMR kinships 
<!--
(i.e., the counts $\KHSS{c_i,c_j}{g_i,s_i}$, $\KHSD{c_i,c_j}{g_i,s_i}$,
$\KPOP{c_i,c_j}{g_i,s_i}{=}$, $\KPOP{c_i,c_j}{g_i,s_i}{\neq}$, and
$\KNot{c_i,c_j}{g_i,s_i}$), 
-->
(i.e., counts of the number of observed HSPs and POPs sharing or not sharing mtDNA
as well as the number of unrelated pairs),
a recruitment (age-1) relative abundance index, and age compositions. 
These served as data inputs for the estimation model (see below). 
Note that, unlike traditional fishery models, we did
not simulate or fit to catch data; scale for absolute abundance estimates was provided solely by
the CKMR data.

Samples for CKMR were drawn
during the final years of the simulation, from all dead individuals in the population, 
sampling age 1 for juveniles and ages 4+ for adults. In reality, adult samples would be obtained
through a combination of fishery-independent and fishery-dependent sampling, as most of the
males would need to be sampled in marine protected areas, given their scarcity in open areas [@lowerre2020testing].
The kin relationship between every pair of sampled
individuals was identified from the pedigree, and identity by descent of mtDNA between relatives
was recorded from the sex of the parent (for POPs) and of the shared parent (for HSPs).  Pairs that did not share mtDNA identical-by-descent were assigned matching
mtDNA with probability $1-H_\mathrm{mt}$. These observations were tallied according
to both individuals' birth years as well as the first-born's sampling year and sex at time of sampling.
Only pairs of samples with different birth years ("cross-cohort
comparisons") were included in the CKMR likelihood to minimize the effect of
interannual and inter-individual variation in the number of offspring [@waples2022close].
The total number of CKMR samples collected across all sampling years (see below) was chosen as various multiples of the
$10\sqrt{N}$ guideline from @bravington2016close for a single sample of a
roughly equal mix of juveniles and adults. We used the average simulated abundance of
age 2+ individuals over the CKMR sampling period as $N$ in the total sample
size calculation and the average sex ratio using all ages to partition samples between males and females.
The effect of increasing or decreasing the relative proportion of males sampled is explored in the supplement.
Thus, the sample sizes varied across simulation replicates of the OM.
The total genetic samples
were spread out over the CKMR sampling period, such that if there were fewer
sampling years, more samples per year would be collected. For each of the 40
simulated pedigrees per OM configuration, dead individuals were subsampled with a different random seed ten 
times to create ten distinct CKMR data sets per pedigree, for a total of 400 CKMR data sets per
OM configuration (the subsampling effect is explored in the Supplement).

@@

### Genetic Sampling Scenarios

@@

## Estimation Model

We employed a standard age-structured deterministic fishery population dynamics
model and integrated the CKMR estimators derived in the Appendix to see how well the 
simulated sex-specific abundances could be estimated, given the data generated from the OM. 
Assumed known were all life history and fishery
vulnerability age schedules except sex change probability, $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$,
and age-specific male reproductive value, $\mathrm{RV}_{a,M}$. The inflection point of the two-parameter 
cumulative normal distribution $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
function was always assumed known but the standard
deviation, $\mathrm{STF}_{SD}$, which controls the function's shape was estimated in some EM runs. 
The exponent of the $\mathrm{RV}$
power function was assumed known in the EM only for females 
($\mathrm{RV}_F^\mathrm{OM} = \mathrm{RV}_F^\mathrm{EM}$, more on this below). Additional estimated parameters were mean
recruitment $\rbar$ and fleet-specific fishing mortality rates $F$ as well as annual
deviations for each. No stock-recruitment relationship was simulated or attempted
to be estimated, and unfished abundance was also not estimated since the focus of
the study was to estimate scale (i.e., absolute abundance) and sex ratio in recent years. Full equations
for the total objective function and its components are presented in Appendix B. 
The estimation model was implemented in AD Model Builder [@fournier2012ad].

### EM Scenarios

The EM for the base case included the full CKMR pseudolikelihood (for both HSPs and POPs)
and assumed that the degree of mitochondrial haplotype diversity in the population ($\Hmt$),
$\tprob_{a}$, and $\mathrm{RV}_M$ were known. 
Alternative scenarios tested models that included HSPs only in the likelihood or
contained misspecification (Figure \@ref(fig:simoverview-plot)). 
One misspecification was an incorrect $\Hmt$, which affects our ability
to correctly infer the sex of an unobserved parent or of an observed parent at a time
prior to sampling, potentially biasing CKMR kinship probabilities.
Other misspecifications were that skipped spawning was simulated 
in the OM but not accounted for in the EM and that 
$\mathrm{RV}_M^\mathrm{EM} = \mathrm{RV}_F^\mathrm{EM}$ when 
$\mathrm{RV}_M^\mathrm{OM} \neq \mathrm{RV}_F^\mathrm{OM}$. 
Finally, we allowed the EM to estimate $\mathrm{RV}_M$ to 
assess whether the model could infer this unknown quantity from the CKMR data, 
and to estimate $\tprob_{a}$ to determine if the CKMR data could allow estimating the sex ratio.

@@

### Model Evaluation
For every assessment year $t$ and simulated observed data set $k$, we calculated percent relative error ($\mathrm{PRE}$)
between simulated and estimated sex-specific 
adult abundance and sex ratio as:
\begin{equation}
\mathrm{PRE}_{t,k} = \frac{(\mathrm{EV}_{t,k}-\mathrm{SV}_{t,k})}{\mathrm{SV}_{t,k}}\times 100,
\end{equation}
where $\mathrm{EV}$ = estimated value and $\mathrm{SV}$ = simulated value. For every $k$, 
the median PRE was calculated across years, and these medians were used to calculate median and interquartile
range ($\mathrm{IQR}$) of PRE to compare accuracy and precision of estimates, respectively, by tested scenario. 
We also calculated $\mathrm{PRE}$ for the main model parameters for every $k$. Sex ratio was
defined as the number of males divided by the total number of 3+ year-olds. We include calculation 
and discussion of an additional metric, the normalized root mean squared error, in the Supplement.

# Results

## Simulated Population
At the beginning of the simulation, the median number of adults across all OM
pedigrees was about 690,000 adult individuals (ages 3+) and about 27% of them were male.
In the final year, the median abundance was ~230,000 individuals without 
the productivity-loss feedback loop from sperm limitation
and ~100,000 with sperm limitation, and about 3% of them were male. When sperm limitation
was present the population trajectory started diverging from the scenarios without
sperm limitation in the final 35-40 years of the simulation (Figure \@ref(fig:simpop-plot)).

## Simulated CKMR Samples
The number of females sampled was about two magnitudes higher than the number of
males sampled (Figure S3). Additionally, slightly fewer overall 
female CKMR samples were collected when CKMR sampling was spread out over the longer period.
This is because 10 years of sampling included years of lower abundance that lowered
the average abundance on which CKMR sample size was based. Conversely, the number
of males sampled tended to be slightly higher for the 10-year sampling time period, because
the sex ratio used to partition sampling by sex was higher for the longer time period.
The median number of males sampled for the base OM scenario ranged from 15
individuals ($\mathrm{IQR} = 3$) when CKMR sample size was 50% of $n_\mathrm{base}$ and sampling was
conducted over three years to 50 individuals ($\mathrm{IQR} = 10$) when CKMR sample size was
150% of $n_\mathrm{base}$ and samples were collected over ten years.
For adult females, the number sampled ranged from a median of 1480 ($\mathrm{IQR} = 100$) for
50% of $n_\mathrm{base}$ and ten sampling years to a median of 4664 ($\mathrm{IQR} = 497$) for
150% of $n_\mathrm{base}$ and three sampling years (Figure S3).

@@

Some differences in the number of pairs found were apparent across OM scenarios
(Figure \@ref(fig:mainckmrpairs-plot)). First, the number of unrelated pairs was much
smaller for skipped spawning than the other scenarios while
the number of related pairs was not. This is because abundance was lower for this
scenario and so fewer individuals needed to be sampled to obtain similar numbers of
related pairs. Additionally, the number of $\HSPS$
was higher for the skipped spawning scenario because younger females
were more likely to skip spawning than older ones, meaning comparatively fewer
females contributed to the offspring produced and the likelihood of finding
HSPs sharing a mother increased.
The next major effect was that of $\mathrm{RV}_M^\mathrm{OM}$ on $\HSPD$.
When $\mathrm{RV}_M^\mathrm{OM}$ increased more steeply with age, as would be
expected based on the size advantage model, the number of $\HSPD$ more than doubled
compared to the base case.
The explanation is analogous to finding more $\HSPS$ when there is skipped spawning. 
If $\mathrm{RV}_M^\mathrm{OM}$ is steeper,
older males (of which there are fewer than younger males) are much more successful at fathering offspring and
hence the chance of finding $\HSPD$ increases.
The opposite is seen when $\mathrm{RV}_M^\mathrm{OM}$ varies little with age. Here,
we only found about two-thirds the number of $\HSPD$ compared to the base case
(Figure \@ref(fig:mainckmrpairs-plot))
because now younger males are also very successful and the offspring are more
evenly spread out among fathers, reducing the likelihood of encountering HSPs sharing
a father.

## Estimation Model Convergence

@@

## Effect of CKMR Information Availability on Estimation Performance

@@

### CKMR sample size and number of sampling years

When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M$ were assumed known, median abundance estimates
had no (median $\mathrm{PRE} < \pm 1\%$) or low (median $\mathrm{PRE} < \pm 2\%$) bias, with no difference between
male and female estimates. Precision was relatively high, with $\mathrm{IQR_{PRE}}$
ranging from 5% to 13% across all tested CKMR sample sizes and numbers of CKMR sampling 
years (Figure \@ref(fig:ckmrsampling-plot)). There was no difference
in median bias between three and ten years of CKMR sampling and precision
was only slightly lower for ten years of sampling. Increasing the CKMR sample size increased
precision in $\mathrm{PRE}$ by a relatively small amount (28%) when sample
size was increased from 100% to 150% of $n_\mathrm{base}$ and more significantly
(${\approx}67\%$) when the sample size increased from 50% to 100% of $n_\mathrm{base}$.

>>>>

bibliography      : ["citations_abbrev.bib", "r-references.bib"]
csl               : "elsevier-harvard.csl"

@@

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
library(gridExtra)
library(dplyr)
library(here)

# caching for quick knitting.  
#    set cache = FALSE to not cache
#    set cache.extra to a new value to remove all caches and re-evaluate
knitr::opts_chunk$set(cache.extra = 3, cache = FALSE)

@@

```{r tables_figs, include = FALSE}
source(here("code", "analysis", "results_tables.R"), local = knitr::knit_global())
source(here("code", "analysis", "results_figures.R"), local = knitr::knit_global())
```

@@

# 1. Introduction

@@

# 2. Methods

This section is organized as follows: 
we first describe the CKMR kinship probabilities for protogynous hermaphrodites, then
present the different components of the simulation study to test this model: the (individual-based) 
operating model (OM) to simulate 'reality', the sampling model to generate 'observed' data from 
the OM, and finally the integrated fisheries stock assessment estimation model (EM) that uses
the 'observed' data and the derived CKMR kinship probabilities to estimate a set of parameters 
that allow generating 'estimated' abundance which can then be compared to the 'true' abundance 
generated by the OM. For each simulation component, we describe the configurations or scenarios 
we used to test the performance of our CKMR estimators (Figure 1).

## 2.1. CKMR Model

@@

The ability to identify---or at least infer---the sex of a sampled individual's parent is crucial for 
applying the model, and in our model this is accomplished by sequencing the fish's
mitochondrial DNA (mtDNA) which is inherited from the female parent. For the purposes
of the simulation analyses, we assumed that different mtDNA haplotypes were equifrequent,
so that we were able to parameterize the mtDNA haplotype frequencies simply by $H_\mathrm{mt}$---the
probability that two randomly sampled mtDNAs from the population
are different in sequence. This quantity is also commonly called the mtDNA
haplotype diversity [@avise2012molecular], and parameterizing the simulations in terms of it allowed
us to avoid simulating mtDNA sequences
in all simulated individuals and also simplified the CKMR likelihood used to analyze the simulations as described
in the Supplement. An $\Hmt$ of 1 would indicate a 100% certainty
that two HSPs with the same mtDNA sequence shared a mother, while any value <1 means there is
a chance (which equals $1 - \Hmt$) that their mtDNA sequence is the same even though they did not
share a mother. We incorporate uncertainty due to $\Hmt$ into our CKMR model
equations (Equations A8, and A14-A17) and in the simulation, we explore the effects of 
misspecifying both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\Hmt$ in the estimation model (Section 2.5.1).

## 2.2. Model Species

@@

## 2.3. Operating Model

The OM is a discrete-time, individual-based model (IBM) implemented in R based on the 'fishSim' 
package [@R-fishSim]. We modified the package functions to accommodate 
our study needs (e.g., by adding the ability to model protogynous hermaphrodite 
life histories) and rewrote the slower-running R functions in C++ using the 'Rcpp' 
package [@eddelbuettel2013seamless] to increase computational efficiency. The basic
annual sequence, modeled loosely on gag life history,
was that sex transition and then spawning occurred at the beginning
of the year, followed by mortality, CKMR sampling, and age incrementation (Figure 2).
For computational efficiency, the simulation starts at age 1, given that most age-0 gag
are not available to the fishery and suffer high rates of natural mortality.

Simulated reproductive value for females and males of age $a$, denoted as  $\mathrm{RV}_{a,F}$ and
$\mathrm{RV}_{a,M}$, respectively, were used to produce age-1 offspring and assign parentage.
$\mathrm{RV}_{a,F}$ represents the age-specific expected number 
of age-1 offspring produced per female. This was modeled as a power function 
(Figure 3, panel A) derived 
from female gag batch fecundity-at-age data, and scaled to achieve a stable population in 
the absence of fishing. To help produce recruitment variability, $\mathrm{RV}_{a,F}$ 
was randomly scaled up or down annually by sampling a reproductive value, 
$\mathrm{RV}_{a,t,F,b}$ for each breeding group, $b$, from a $\mathrm{Uniform}(0,1)$ 
distribution and rescaling such that $\sum_a{\mathrm{RV}_{a,F}e^{\eta_t}} = \sum_a{\sum_b\mathrm{RV}_{a,t,F,b}}$.
Here, $\eta_t$ represents the annual scaling factor, ensuring that the total RV 
across all ages and breeding groups remains consistent.
The realized number of age-1 offspring in a given year $t$
for a given female of age $a$ in breeding group $b$ was a Poisson random draw from 
$\mathrm{RV}_{a,t,F,b}$. All males were considered mature, and $\mathrm{RV}_M$ 
was used as the age-specific probability of being the father
to any age-1 offspring produced by females in the male's breeding group. The true shape and 
parameter values of $\mathrm{RV}_M$ are unknown. Here, we modeled it as a power function, and in 
the base OM its exponent was set equal to that of $\mathrm{RV}_F$ (Figure 3, panel A), 
which corresponds to assuming age-specific $\mathrm{RV}_M$ equals that of the females (the current
assumption of the gag stock assessment). Mating was assumed to 
be fully random within breeding groups.

@@

The simulation was run for 119 years, starting with an unfished population and 
ending with a highly depleted population. For each OM scenario (Section 2.3.1), 40
independent replicates of the population were simulated, providing 40
unique whole-population pedigrees.

### 2.3.1. OM Scenarios

We simulated four alternative versions of 'reality' in our OM to explore the
effect of uncertainty about male contribution to reproductive success: in addition to
the base case, one scenario included sperm limitation and two scenarios
simulated (age-specific) male reproductive value functions ($\mathrm{RV}_M^\mathrm{OM}$)
that differed from those of the females ($\mathrm{RV}_M^\mathrm{EM}$). In the base OM, male 
and female reproductive value were equal, corresponding 
to what is currently assumed in the gag stock assessment, but in two alternative 
scenarios, we specified $\mathrm{RV}_M^\mathrm{OM}$ functions that increased more or less steeply
from $\mathrm{RV}_F^\mathrm{OM}$ (Figure 3, panel A).

The true $\mathrm{RV}_M$ in protogynous hermaphrodites, including gag,
is unknown. It is assumed that sperm limitation exists but it has not been
measured. However, it has been shown that skipped spawning is relatively common in female gag,
and this is thought to be linked to low male abundance [@lowerre2022spawning].
In the base OM configuration, declining male abundance had no effect (meaning all eggs produced
by females will get fertilized, regardless of sex ratio). Alternately, we simulated the effect of
sperm limitation as an increased
likelihood of a female not producing any offspring in a given spawning season as the sex ratio
declines. Note that a female skipping spawning was independent between years and thus there was no systematic
skipped spawning such as alternate year spawning. The relationship between the 
current proportion of males in a spawning group relative to unfished levels, $p_M$,
and the probability of females in that spawning group skipping spawning was
modeled as an exponential function of the form $1-\exp(-{\tau p_M})$ (Figure
3, panel B), where $\tau$ controls the slope of the curve. We used a sigmoid
function to map probability of spawning to age, such that skipped spawning was more likely
in younger females. Thus, skipped spawning serves as a de facto modifier of the vector 
describing the age-specific probability of a female producing offspring in a given year.

## 2.4. Sampling Model
Three types of observations were generated from the OM for use in the
integrated assessment model: CKMR kinships 
<!--
(i.e., the counts $\KHSS{c_i,c_j}{g_i,s_i}$, $\KHSD{c_i,c_j}{g_i,s_i}$,
$\KPOP{c_i,c_j}{g_i,s_i}{=}$, $\KPOP{c_i,c_j}{g_i,s_i}{\neq}$, and
$\KNot{c_i,c_j}{g_i,s_i}$), 
-->
(i.e., counts of the number of observed HSPs and POPs sharing or not sharing mtDNA
as well as the number of unrelated pairs),
a recruitment (age-1) relative abundance index, and age compositions. 
These served as data inputs for the estimation model (Section 2.5). 
Note that, unlike traditional fishery models, we did
not simulate or fit to catch data; scale for absolute abundance estimates was provided solely by
the CKMR data.

Samples for CKMR were drawn
during the final years of the simulation, from all dead individuals in the population, 
sampling age 1 for juveniles and ages 4+ for adults. In reality, adult samples would be obtained
through a combination of fishery-independent and fishery-dependent sampling, as most of the
males would need to be sampled in marine protected areas, given their scarcity in open areas [@lowerre2020testing].
The kin relationship between every pair of sampled
individuals was identified from the pedigree, and identity by descent of mtDNA between relatives
was recorded from the sex of the parent (for POPs) and of the shared parent (for HSPs).  Pairs that did not share mtDNA identical-by-descent were assigned matching
mtDNA with probability $1-H_\mathrm{mt}$. These observations were tallied according
to both individuals' birth years as well as the first-born's sampling year and sex at time of sampling.
Only pairs of samples with different birth years ("cross-cohort
comparisons") were included in the CKMR likelihood to minimize the effect of
interannual and inter-individual variation in the number of offspring [@waples2022close].
The total number of CKMR samples collected across all sampling years (Section 2.4.1) was chosen as various multiples of the
$10\sqrt{N}$ guideline from @bravington2016close for a single sample of a
roughly equal mix of juveniles and adults. We used the average simulated abundance of
age 2+ individuals over the CKMR sampling period as $N$ in the total sample
size calculation and the average sex ratio using all ages to partition samples between males and females.
The effect of increasing or decreasing the relative proportion of males sampled is explored in the supplement.
Thus, the sample sizes varied across simulation replicates of the OM.
The total genetic samples
were spread out over the CKMR sampling period, such that if there were fewer
sampling years, more samples per year would be collected. For each of the 40
simulated pedigrees per OM configuration, dead individuals were subsampled with a different random seed ten 
times to create ten distinct CKMR data sets per pedigree, for a total of 400 CKMR data sets per
OM configuration (the subsampling effect is explored in the Supplement).

@@

### 2.4.1. Genetic Sampling Scenarios

@@

## 2.5. Estimation Model

We employed a standard age-structured deterministic fishery population dynamics
model and integrated the CKMR estimators derived in the Appendix to see how well the 
simulated sex-specific abundances could be estimated, given the data generated from the OM. 
Assumed known were all life history and fishery
vulnerability age schedules except sex change probability, $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$,
and age-specific male reproductive value, $\mathrm{RV}_{a,M}$. The inflection point of the two-parameter 
cumulative normal distribution $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
function was always assumed known but the standard
deviation, $\mathrm{STF}_{SD}$, which controls the function's shape was estimated in some EM runs. 
The exponent of the $\mathrm{RV}$
power function was assumed known in the EM only for females 
($\mathrm{RV}_F^\mathrm{OM} = \mathrm{RV}_F^\mathrm{EM}$, more on this in Section 2.5.1). Additional estimated parameters were mean
recruitment $\rbar$ and fleet-specific fishing mortality rates $F$ as well as annual
deviations for each. No stock-recruitment relationship was simulated or attempted
to be estimated, and unfished abundance was also not estimated since the focus of
the study was to estimate scale (i.e., absolute abundance) and sex ratio in recent years. Full equations
for the total objective function and its components are presented in Appendix B. 
The estimation model was implemented in AD Model Builder [@fournier2012ad].

### 2.5.1. EM Scenarios

The EM for the base case included the full CKMR pseudolikelihood (for both HSPs and POPs)
and assumed that the degree of mitochondrial haplotype diversity in the population ($\Hmt$),
$\tprob_{a}$, and $\mathrm{RV}_M$ were known. 
Alternative scenarios tested models that included HSPs only in the likelihood or
contained misspecification (Figure 1). 
One misspecification was an incorrect $\Hmt$, which affects our ability
to correctly infer the sex of an unobserved parent or of an observed parent at a time
prior to sampling, potentially biasing CKMR kinship probabilities.
Other misspecifications were that skipped spawning was simulated 
in the OM but not accounted for in the EM and that 
$\mathrm{RV}_M^\mathrm{EM} = \mathrm{RV}_F^\mathrm{EM}$ when 
$\mathrm{RV}_M^\mathrm{OM} \neq \mathrm{RV}_F^\mathrm{OM}$. 
Finally, we allowed the EM to estimate $\mathrm{RV}_M$ to 
assess whether the model could infer this unknown quantity from the CKMR data, 
and to estimate $\tprob_{a}$ to determine if the CKMR data could allow estimating the sex ratio.

@@

### 2.5.2.  Model Evaluation
For every assessment year $t$ and simulated observed data set $k$, we calculated percent relative error ($\mathrm{PRE}$)
between simulated and estimated sex-specific 
adult abundance and sex ratio as:
\begin{equation}
\mathrm{PRE}_{t,k} = \frac{(\mathrm{EV}_{t,k}-\mathrm{SV}_{t,k})}{\mathrm{SV}_{t,k}}\times 100,
\end{equation}
where $\mathrm{EV}$ = estimated value and $\mathrm{SV}$ = simulated value. For every $k$, 
the median PRE was calculated across years, and these medians were used to calculate median and interquartile
range ($\mathrm{IQR}$) of PRE to compare accuracy and precision of estimates, respectively, by tested scenario. 
We also calculated $\mathrm{PRE}$ for the main model parameters for every $k$. Sex ratio was
defined as the number of males divided by the total number of 3+ year-olds. We include calculation 
and discussion of an additional metric, the normalized root mean squared error, in the Supplement.

# 3. Results

## 3.1. Simulated Population
At the beginning of the simulation, the median number of adults across all OM
pedigrees was about 690,000 adult individuals (ages 3+) and about 27% of them were male.
In the final year, the median abundance was ~230,000 individuals without 
the productivity-loss feedback loop from sperm limitation
and ~100,000 with sperm limitation, and about 3% of them were male. When sperm limitation
was present the population trajectory started diverging from the scenarios without
sperm limitation in the final 35-40 years of the simulation (Figure 4).

## 3.2. Simulated CKMR Samples
The number of females sampled was about two magnitudes higher than the number of
males sampled (Figure S3). Additionally, slightly fewer overall 
female CKMR samples were collected when CKMR sampling was spread out over the longer period.
This is because 10 years of sampling included years of lower abundance that lowered
the average abundance on which CKMR sample size was based. Conversely, the number
of males sampled tended to be slightly higher for the 10-year sampling time period, because
the sex ratio used to partition sampling by sex was higher for the longer time period.
The median number of males sampled for the base OM scenario ranged from 15
individuals ($\mathrm{IQR} = 3$) when CKMR sample size was 50% of $n_\mathrm{base}$ and sampling was
conducted over three years to 50 individuals ($\mathrm{IQR} = 10$) when CKMR sample size was
150% of $n_\mathrm{base}$ and samples were collected over ten years.
For adult females, the number sampled ranged from a median of 1480 ($\mathrm{IQR} = 100$) for
50% of $n_\mathrm{base}$ and ten sampling years to a median of 4664 ($\mathrm{IQR} = 497$) for
150% of $n_\mathrm{base}$ and three sampling years (Figure S3).

@@

Some differences in the number of pairs found were apparent across OM scenarios
(Figure 5). First, the number of unrelated pairs was much
smaller for skipped spawning than the other scenarios while
the number of related pairs was not. This is because abundance was lower for this
scenario and so fewer individuals needed to be sampled to obtain similar numbers of
related pairs. Additionally, the number of $\HSPS$
was higher for the skipped spawning scenario because younger females
were more likely to skip spawning than older ones, meaning comparatively fewer
females contributed to the offspring produced and the likelihood of finding
HSPs sharing a mother increased.
The next major effect was that of $\mathrm{RV}_M^\mathrm{OM}$ on $\HSPD$.
When $\mathrm{RV}_M^\mathrm{OM}$ increased more steeply with age, as would be
expected based on the size advantage model, the number of $\HSPD$ more than doubled
compared to the base case.
The explanation is analogous to finding more $\HSPS$ when there is skipped spawning. 
If $\mathrm{RV}_M^\mathrm{OM}$ is steeper,
older males (of which there are fewer than younger males) are much more successful at fathering offspring and
hence the chance of finding $\HSPD$ increases.
The opposite is seen when $\mathrm{RV}_M^\mathrm{OM}$ varies little with age. Here,
we only found about two-thirds the number of $\HSPD$ compared to the base case
(Figure 5)
because now younger males are also very successful and the offspring are more
evenly spread out among fathers, reducing the likelihood of encountering HSPs sharing
a father.

## 3.3. Estimation Model Convergence

@@

## 3.4. Effect of CKMR Information Availability on Estimation Performance

@@

### 3.4.1. CKMR sample size and number of sampling years

When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M$ were assumed known, median abundance estimates
had no (median $\mathrm{PRE} < \pm 1\%$) or low (median $\mathrm{PRE} < \pm 2\%$) bias, with no difference between
male and female estimates. Precision was relatively high, with $\mathrm{IQR_{PRE}}$
ranging from 5% to 13% across all tested CKMR sample sizes and numbers of CKMR sampling 
years (Figure 6). There was no difference
in median bias between three and ten years of CKMR sampling and precision
was only slightly lower for ten years of sampling. Increasing the CKMR sample size increased
precision in $\mathrm{PRE}$ by a relatively small amount (28%) when sample
size was increased from 100% to 150% of $n_\mathrm{base}$ and more significantly
(${\approx}67\%$) when the sample size increased from 50% to 100% of $n_\mathrm{base}$.
